The Proofs

### 2023-11-13 proofs/tex/decl-subtyping-lemmas.tex
Restructure proofs

\begin{lemma}[Free Variable Propagation] \label{lemma:fv-propagation}
  In the judgments of negative subtyping or positive supertyping,
  free variables propagate left-to-right. For a context $[[Γ]]$,
  \begin{itemize}
    \item $-$ if $[[Γ ⊢ iN ≤ iM]]$ then $[[fv(iN)]] \subseteq [[fv(iM)]]$
    \item $+$ if $[[Γ ⊢ iP ≥ iQ]]$ then $[[fv(iP)]] \subseteq [[fv(iQ)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  Mutual induction on $[[Γ ⊢ iN ≤ iM]]$ and $[[Γ ⊢ iP ≥ iQ]]$.
  \begin{caseof}
  \item $[[G ⊢ a⁻ ≤ a⁻]]$\\
    It is self-evident that $[[{a⁻} ⊆ {a⁻}]]$.
  \item $[[G ⊢ ↑iP ≤ ↑iQ]]$
    From the inversion (and unfolding $[[G ⊢ iP ≈ iQ]]$ ), we have
    $[[G ⊢ iP ≥ iQ]]$. Then by the induction hypothesis,
    $[[fv(iP)]] \subseteq [[fv(iQ)]]$. The desired 
    inclusion holds, since $[[fv(↑iP)]] = [[fv(iP)]]$ and
    $[[fv(↑iQ)]] = [[fv(iQ)]]$.
  \item $[[G ⊢ iP → iN ≤ iQ → iM]]$
    The induction hypothesis applied to the premises gives:
    $[[fv(iP)]] \subseteq [[fv(iQ)]]$ and
    $[[fv(iN)]] \subseteq [[fv(iM)]]$.
    Then $[[fv(iP → iN)]] = [[fv(iP) ∪ fv(iN)]] \subseteq
    [[fv(iQ) ∪ fv(iM)]] = [[fv(iQ → iM)]]$.

@@

\begin{corollary}[Free Variables of mutual subtypes] \label{corollary:fv-mut-sub}
  \hfill
  \begin{itemize}
    \item [$-$] If $[[Γ ⊢ iN ≈ iM]]$ then $[[fv iN]] = [[fv iM]]$, 
    \item [$+$] If $[[Γ ⊢ iP ≈ iQ]]$ then $[[fv iP]] = [[fv iQ]]$
  \end{itemize}
\end{corollary}

\begin{lemma}[Decomposition of quantifier rules]
  \label{lemma:quant-rule-decomposition}
  Assuming that $[[pas]]$, $[[pbs]]$, $[[nas]]$, and $[[nas]]$ are disjoint from $[[Γ]]$,
  \begin{itemize}
    \item [$-_{R}$] $[[Γ ⊢ iN ≤ ∀pbs.iM]]$ holds if and only if $[[Γ, pbs ⊢ iN ≤ iM]]$;
    \item [$+_{R}$] $[[Γ ⊢ iP ≥ ∃nbs.iQ]]$ holds if and only if $[[Γ, nbs ⊢ iP ≥ iQ]]$;
    \item [$-_{L}$] suppose $[[iM]] \neq [[∀]]\dots$
      then $[[Γ ⊢ ∀pas.iN ≤ iM]]$ holds if and only if $[[Γ ⊢ [iPs/pas]iN ≤ iM]]$
      for some $[[Γ ⊢ iPs]]$;
    \item [$+_{L}$] suppose $[[iQ]] \neq [[∃]]\dots$
      then $[[Γ ⊢ ∃nas.iP ≥ iQ]]$ holds if and only if $[[Γ ⊢ [iNs/nas]iP ≥ iQ]]$
      for some $[[Γ ⊢ iNs]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  \hfill
  \begin{itemize}
    \item [$-_{R}$] Let us prove both directions. 
      \begin{itemize}
        \item [$\Rightarrow$] Let us assume $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.
          $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.
          Let us decompose $[[iM]]$ as $[[∀pbs'.iM']]$ where $[[iM']]$ does not start with $[[∀]]$, 
          and decompose $[[iN]]$ as $[[∀pas.iN']]$ where $[[iN']]$ does not start with $[[∀]]$.
          If $[[pbs]]$ is empty, then $[[Γ, pbs ⊢ iN ≤ iM]]$ holds by assumption.
          Otherwise, $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs.∀pbs'.iM]]$ is inferred by
          \ruleref{\ottdruleDOneForallLabel}, and by inversion:
          $[[Γ, {pbs}, {pbs'} ⊢ [iPs/pas]iN' ≤ iM']]$ for some $[[Γ, {pbs}, {pbs'} ⊢ iPs]]$.
          Then again by \ruleref{\ottdruleDOneForallLabel} with the same $[[iPs]]$,
          $[[Γ,pbs ⊢ ∀pas.iN' ≤ ∀pbs'.iM']]$, that is $[[Γ,pbs ⊢ iN ≤ iM]]$.
        \item [$\Leftarrow$] let us assume $[[Γ, pbs ⊢ iN ≤ iM]]$, and let us decompose 
          $[[iN]]$ as $[[∀pas.iN']]$ where $[[iN']]$ does not start with $[[∀]]$, 
          and $[[iM]]$ as $[[∀pbs'.iM']]$ where $[[iM']]$ does not start with $[[∀]]$.
          if $[[pas]]$ and $[[pbs']]$ are empty then $[[Γ, pbs ⊢ iN ≤ iM]]$
          is turned into $[[Γ ⊢ iN ≤ ∀pbs.iM]]$ by \ruleref{\ottdruleDOneForallLabel}.
          Otherwise, $[[Γ, pbs ⊢ ∀pas.iN' ≤ ∀pbs'.iM']]$ is inferred by
          \ruleref{\ottdruleDOneForallLabel}, that is $[[Γ, {pbs}, {pbs'} ⊢ [iPs/pas]iN' ≤ iM']]$
          for some $[[Γ, {pbs}, {pbs'} ⊢ iPs]]$.
          Then by \ruleref{\ottdruleDOneForallLabel} again,
          $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs,pbs'.iM']]$, in other words, $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs.∀pbs'.iM']]$, 
          that is $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.

@@

\begin{corollary}[Redundant quantifier elimination]
  \label{corollary:red-quant-elim}
  \hfill
  \begin{itemize}
    \item [$-_{L}$] Suppose that $[[ {pas} ∩ fv(iN) = ∅]]$ then 
      $[[Γ ⊢ ∀pas.iN ≤ iM]]$ holds if and only if $[[Γ ⊢ iN ≤ iM]]$;
    \item [$-_{R}$] Suppose that $[[ {pas} ∩ fv(iM) = ∅]]$ then 
      $[[Γ ⊢ iN ≤ ∀pas.iM]]$ holds if and only if $[[Γ ⊢ iN ≤ iM]]$;
    \item [$+_{L}$] Suppose that $[[ {nas} ∩ fv(iP) = ∅]]$ then
      $[[Γ ⊢ ∃nas.iP ≥ iQ]]$ holds if and only if $[[Γ ⊢ iP ≥ iQ]]$.
    \item [$+_{R}$] Suppose that $[[ {nas} ∩ fv(iQ) = ∅]]$ then 
      $[[Γ ⊢ iP ≥ ∃nas.iQ]]$ holds if and only if $[[Γ ⊢ iP ≥ iQ]]$.
  \end{itemize}
\end{corollary}
\begin{proof}
  \begin{itemize}
    \item [$-_{R}$] Suppose that $[[ {pas} ∩ fv(iM) = ∅]]$ then 
      by \cref{lemma:quant-rule-decomposition},
      $[[Γ ⊢ iN ≤ ∀pas.iM]]$ 
      is equivalent to $[[Γ, pas ⊢ iN ≤ iM]]$,
      By \cref{lemma:wf-ctxt-equiv},
      since $[[{pas} ∩ fv(iN) = ∅]]$ and $[[{pas} ∩ fv(iM) = ∅]]$,
      $[[Γ, pas ⊢ iN ≤ iM]]$ is equivalent to $[[Γ ⊢ iN ≤ iM]]$.

@@

\begin{lemma}[Subtypes and supertypes of a variable]
  \label{lemma:var-subt}
  Assuming $[[Γ ⊢  α⁻]]$, $[[Γ ⊢ α⁺]]$, $[[Γ ⊢ iN]]$, and $[[Γ ⊢ iP]]$,
  \begin{itemize}
  \item[$+$] if $[[Γ ⊢ iP ≥ ∃nas.α⁺]]$ or $[[Γ ⊢ ∃nas.α⁺ ≥ iP ]]$ then $[[iP]] = [[∃nbs.α⁺]]$ (for some potentially empty $[[nbs]]$)
  \item[$-$] if $[[Γ ⊢ iN ≤ ∀pas.α⁻]]$ or $[[Γ ⊢ ∀pas.α⁻ ≤ iN ]]$ then $[[iN]] = [[∀pbs.α⁻]]$ (for some potentially empty $[[pbs]]$)
  \end{itemize}
\end{lemma}
\begin{proof}
  We prove by induction on the tree
  inferring $[[Γ ⊢ iP ≥ ∃nas.α⁺]]$ or $[[Γ ⊢ ∃nas.α⁺ ≥ iP ]]$ or
  or $[[Γ ⊢ iN ≤ ∀pas.α⁻]]$ or $[[Γ ⊢ ∀pas.α⁻ ≤ iN ]]$.

@@

\begin{corollary}[Variables have no proper subtypes and supertypes]
  \label{corollary:vars-no-proper-subtypes}
  Assuming that all mentioned types are well-formed in $[[Γ]]$,
  \begin{align*}
    [[Γ ⊢ iP ≥ α⁺]] ~ &\iff ~ [[iP = ∃nbs.α⁺]]  ~ \iff ~ [[Γ ⊢ iP ≈ α⁺]] ~ \iff ~ [[iP ≈ α⁺]]\\
    [[Γ ⊢ α⁺≥ iP]]  ~ &\iff ~ [[iP = ∃nbs.α⁺]]  ~ \iff ~ [[Γ ⊢ iP ≈ α⁺]] ~ \iff ~ [[iP ≈ α⁺]]\\
    [[Γ ⊢ iN ≤ α⁻]] ~ &\iff ~ [[iN = ∀pbs.α⁻]]  ~ \iff ~ [[Γ ⊢ iN ≈ α⁻]] ~ \iff ~ [[iN ≈ α⁻]]\\
    [[Γ ⊢ α⁻ ≤ iN]] ~ &\iff ~ [[iN = ∀pbs.α⁻]]  ~ \iff ~ [[Γ ⊢ iN ≈ α⁻]] ~ \iff ~ [[iN ≈ α⁻]]\\
  \end{align*}
\end{corollary}
\begin{proof}
  Notice that $[[Γ ⊢ ∃nbs.α⁺ ≈ α⁺]]$ and $[[∃nbs.α⁺ ≈ α⁺]]$ and apply
  \cref{lemma:var-subt}.
\end{proof}

\begin{lemma}[Subtyping context irrelevance]
  \label{lemma:subt-ctxt-irrelevance}
  Suppose that all the mentioned types are well-formed in $[[Γ1]]$ and $[[Γ2]]$. 
  Then 
  \begin{enumerate}
    \item[$+$] $[[Γ1 ⊢ iP ≥ iQ]]$ is equivalent to $[[Γ2 ⊢ iP ≥ iQ]]$;
    \item[$-$] $[[Γ1 ⊢ iN ≤ iM]]$ is equivalent to $[[Γ2 ⊢ iN ≤ iM]]$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  We prove it by induction on 
  the size of $[[Γ1 ⊢ iP ≥ iQ]]$ and mutually,
  the size of $[[Γ1 ⊢ iN ≤ iM]]$.

@@

  Suppose that the judgement is inferred by \ruleref{\ottdruleDOneExistsLabel}.
  Then we are proving that $[[Γ1 ⊢ ∃nas.iP ≥ ∃nbs.iQ]]$ implies 
  $[[Γ2 ⊢ ∃nas.iP ≥ ∃nbs.iQ]]$ (the other implication is proven symmetrically).

@@

  Notice that the premises of the induction hold:
  \begin{enumerate}
    \item $[[Γi, nbs ⊢ [σ0]iP]]$, 
    \item $[[Γi, nbs ⊢ iQ]]$, and
    \item $[[Γ1, nbs ⊢ [σ0]iP ≥ iQ ]]$, 
      notice that the tree inferring this judgement 
      is the same tree inferring $[[Γ1, nbs ⊢ [σ]iP ≥ iQ ]]$ 
      (since $[[ [σ0]iP = [σ]iP]]$), i.e., 
      it is a subtree of $[[Γ1 ⊢ ∃nas.iP ≥ ∃nbs.iQ]]$.
  \end{enumerate}
  This way, by the induction hypothesis,
  $[[Γ2, nbs ⊢ [σ0]iP ≥ iQ ]]$.
  Combining it with $[[ Γ2, nbs ⊢ σ0 :{nas}]]$
  by \ruleref{\ottdruleDOneExistsLabel},
  we obtain $[[Γ2 ⊢ ∃nas.iP ≥ ∃nbs.iQ]]$.

@@

\begin{lemma}[Weakening of subtyping context]
  \label{lemma:subt-weakening}
  Suppose $[[Γ1]]$ and $[[Γ2]]$ are contexts 
  and $[[ Γ1 ⊆ Γ2 ]]$.
  Then 
  \begin{enumerate}
    \item [$+$] $[[Γ1 ⊢ iP ≥ iQ]]$ implies $[[Γ2 ⊢ iP ≥ iQ]]$;
    \item [$-$] $[[Γ1 ⊢ iN ≤ iM]]$ implies $[[Γ2 ⊢ iN ≤ iM]]$.
  \end{enumerate}
\end{lemma}
\begin{proof}
\end{proof}

\begin{lemma}[Reflexivity of subtyping] \label{lemma:subtyping-reflexivity}
  Assuming all the types are well-formed in $[[Γ]]$,
  \begin{itemize}
    \item [$-$] $[[Γ ⊢ iN ≤ iN]]$
    \item [$+$] $[[Γ ⊢ iP ≥ iP]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  Let us prove it by the size of $[[iN]]$ and mutually, $[[iP]]$.
  \begin{caseof}
    \item $[[iN]] = [[α⁻]]$\\
      Then $[[Γ ⊢ α⁻ ≤ α⁻]]$ is inferred immediately by \ruleref{\ottdruleDOneNVarLabel}.
    \item $[[iN]] = [[∀pas.iN']]$ where $[[pas]]$ is not empty\\
      First, we rename $[[pas]]$ to fresh $[[pbs]]$ in $[[∀pas.iN']]$ to avoid
      name clashes: $[[∀pas.iN']] = [[∀pbs.[pas/pbs]iN']]$.
      Then to infer $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs.[pas/pbs]iN']]$ we can apply 
      \ruleref{\ottdruleDOneForallLabel}, instantiating $[[pas]]$ with $[[pbs]]$:
      \begin{itemize}
        \item $[[fv iN ∩ {pbs} = ∅ ]]$ by choice of $[[pbs]]$,
        \item $[[G, pbs ⊢ pbi]]$,
        \item $[[G, pbs ⊢ [pbs/pas] iN' ≤ [pbs/pas] iN']]$ by the induction hypothesis,
        since the size of $[[ [pbs/pas]iN' ]]$ is equal to the size of $[[iN']]$,
        which is smaller than the size of $[[iN]] = [[∀pas.iN']]$.
      \end{itemize}
    \item $[[iN]] = [[iP → iM]]$\\
      Then $[[Γ ⊢ iP → iM ≤ iP → iM]]$ is inferred by \ruleref{\ottdruleDOneArrowLabel},
      since $[[Γ ⊢ iP ≥ iP]]$ and $[[Γ ⊢ iM ≤ iM]]$ hold the induction hypothesis. 
    \item $[[iN]] = [[↑iP]]$\\
      Then $[[Γ ⊢ ↑iP ≤ ↑iP]]$ is inferred by \ruleref{\ottdruleDOneShiftULabel},
      since $[[Γ ⊢ iP ≥ iP]]$ holds by the induction hypothesis.
    \item The positive cases are symmetric to the negative ones.
  \end{caseof}
\end{proof}

\begin{lemma}[Substitution preserves subtyipng]
  \label{lemma:subst-pres-subt}
  Suppose that all mentioned types are well-formed in $[[Γ1]]$,
  and $[[σ]]$ is a substitution $[[Γ2 ⊢ σ : Γ1]]$.
  \begin{itemize}
    \item [$-$] If $[[Γ1 ⊢ iN ≤ iM]]$ then $[[Γ2 ⊢ [σ]iN ≤ [σ]iM]]$.
    \item [$+$] If $[[Γ1 ⊢ iP ≥ iQ]]$ then $[[Γ2 ⊢ [σ]iP ≥ [σ]iQ]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  We prove it by induction on the size of the derivation of $[[Γ1 ⊢ iN ≤ iM]]$
  and mutually, $[[Γ1 ⊢ iP ≥ iQ]]$. Let us consider the last rule 
  used in the derivation:
  \begin{caseof}
    \item \ruleref{\ottdruleDOneNVarLabel}. Then by inversion, 
      $[[iN = α⁻]]$ and $[[iM = α⁻]]$. By reflexivity of subtyping
      (\cref{lemma:subtyping-reflexivity}),
      we have $[[Γ2 ⊢ [σ]α⁻ ≤ [σ]α⁻]]$, i.e. $[[Γ2 ⊢ [σ]iN ≤ [σ]iM]]$,
      as required.
    \item  \ruleref{\ottdruleDOneForallLabel}. Then by inversion,
      $[[iN = ∀pas.iN']]$, $[[iM = ∀pbs.iM']]$, where $[[pas]]$ or $[[pbs]]$ is not empty.
      Moreover, $[[Γ1, pbs ⊢ [iPs/pas]iN' ≤ iM']]$ for some $[[Γ1, pbs ⊢ iPs]]$, and 
      $[[fv iN ∩ {pbs} = ∅ ]]$.

@@

\begin{corollary}[Substitution preserves subtyping induced equivalence]
  \label{corollary:subst-pres-equiv}
  Suppose that $[[Γ ⊢ σ : Γ1]]$. Then
    \begin{itemize}
      \item[$+$] if $[[Γ1 ⊢ iP]]$,~ $[[Γ1 ⊢ iQ]]$,~ and $[[Γ1 ⊢ iP ≈ iQ]]$ ~ 
        then $[[Γ ⊢ [σ]iP ≈ [σ]iQ]]$
      \item[$-$] if $[[Γ1 ⊢ iN]]$,~ $[[Γ1 ⊢ iM]]$,~ and $[[Γ1 ⊢ iN ≈ iM]]$ ~ 
        then $[[Γ ⊢ [σ]iN ≈ [σ]iM]]$
    \end{itemize}
\end{corollary}

\begin{lemma}[Transitivity of subtyping] 
  \label{lemma:subtyping-transitivity}
  Assuming the types are well-formed in $[[Γ]]$,
  \begin{itemize}
    \item[$-$] if $[[Γ ⊢ iN1 ≤ iN2]]$ and $[[Γ ⊢ iN2 ≤ iN3]]$ then $[[Γ ⊢ iN1 ≤ iN3]]$,
    \item[$+$] if $[[Γ ⊢ iP1 ≥ iP2]]$ and $[[Γ ⊢ iP2 ≥ iP3]]$ then $[[Γ ⊢ iP1 ≥ iP3]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  To prove it, we formulate a stronger property, 
  which will imply the required one, taking $[[σ]] = [[Γ ⊢ id : Γ]]$.

@@

      Notice that the derivations of $[[Γ ⊢ iP ≥ iQ1]]$ and $[[Γ ⊢ iN' ≤ iM1']]$ are proper sub-derivations of
      $[[Γ ⊢ iP → iN' ≤ iQ1 → iM1']]$, and the derivations of $[[Γ ⊢ iQ2 ≥ iR]]$ and $[[Γ ⊢ iM2' ≤ iK']]$ are proper sub-derivations of
      $[[Γ ⊢ iQ2 → iM2' ≤ iR → iK']]$. This way, the induction hypothesis is applicable:
      \begin{itemize}
        \item applying the induction hypothesis to $[[Γ ⊢ iP ≥ iQ1]]$ and $[[Γ ⊢ iQ2 ≥ iR]]$ 
          with $[[Γ' ⊢ σ : Γ]]$ unifying $[[iQ1]]$ and $[[iQ2]]$, we obtain $[[Γ' ⊢ [σ]iP ≥ [σ]iR]]$;
        \item applying the induction hypothesis to $[[Γ ⊢ iN' ≤ iM1']]$ and $[[Γ ⊢ iM2' ≤ iK']]$ 
          with $[[Γ' ⊢ σ : Γ]]$ unifying $[[iM1']]$ and $[[iM2']]$, we obtain $[[Γ' ⊢ [σ]iN' ≤ [σ]iK']]$.
      \end{itemize}
      This way, by \ruleref{\ottdruleDOneArrowLabel}, $[[Γ' ⊢ [σ]iP → [σ]iN' ≤ [σ]iR → [σ]iK']]$,
      that is $[[Γ' ⊢ [σ]iN ≤ [σ]iK]]$, as required.
  \end{caseof}
  After that we consider all the 
  analogous positive cases, and prove them symmetrically.
\end{proof}

\begin{corollary}[Transitivity of equivalence] \label{corollary:equivalence-transitivity}
  Assuming the types are well-formed in $[[Γ]]$,
  \begin{itemize}
    \item[$-$] if $[[Γ ⊢ iN1 ≈ iN2]]$ and $[[Γ ⊢ iN2 ≈ iN3]]$ then $[[Γ ⊢ iN1 ≈ iN3]]$,
    \item[$+$] if $[[Γ ⊢ iP1 ≈ iP2]]$ and $[[Γ ⊢ iP2 ≈ iP3]]$ then $[[Γ ⊢ iP1 ≈ iP3]]$.
  \end{itemize}
\end{corollary}

>>>>

\lemmaFvPropagation*
\begin{proof}
  Mutual induction on $[[Γ ⊢ iN ≤ iM]]$ and $[[Γ ⊢ iP ≥ iQ]]$.
  \begin{caseof}
  \item $[[G ⊢ a⁻ ≤ a⁻]]$\\
    It is self-evident that $[[{a⁻} ⊆ {a⁻}]]$.
  \item $[[G ⊢ ↑iP ≤ ↑iQ]]$
    From the inversion (and unfolding $[[G ⊢ iP ≈ iQ]]$ ), we have
    $[[G ⊢ iP ≥ iQ]]$. Then by the induction hypothesis,
    $[[fv(iP)]] \subseteq [[fv(iQ)]]$. The desired 
    inclusion holds, since $[[fv(↑iP)]] = [[fv(iP)]]$ and
    $[[fv(↑iQ)]] = [[fv(iQ)]]$.
  \item $[[G ⊢ iP → iN ≤ iQ → iM]]$
    The induction hypothesis applied to the premises gives:
    $[[fv(iP)]] \subseteq [[fv(iQ)]]$ and
    $[[fv(iN)]] \subseteq [[fv(iM)]]$.
    Then $[[fv(iP → iN)]] = [[fv(iP) ∪ fv(iN)]] \subseteq
    [[fv(iQ) ∪ fv(iM)]] = [[fv(iQ → iM)]]$.

@@

\corollaryFvMutSub*

\lemmaMutSubTypesWfEquiv*
\begin{proof}
  From \cref{lemma:wf-ctxt-equiv,corollary:fv-mut-sub}.
\end{proof}

\lemmaQuantRuleDecomposition*
\begin{proof}
  \hfill
  \begin{itemize}
    \item [$-_{R}$] Let us prove both directions. 
      \begin{itemize}
        \item [$\Rightarrow$] Let us assume $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.
          $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.
          Let us decompose $[[iM]]$ as $[[∀pbs'.iM']]$ where $[[iM']]$ does not start with $[[∀]]$, 
          and decompose $[[iN]]$ as $[[∀pas.iN']]$ where $[[iN']]$ does not start with $[[∀]]$.
          If $[[pbs]]$ is empty, then $[[Γ, pbs ⊢ iN ≤ iM]]$ holds by assumption.
          Otherwise, $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs.∀pbs'.iM]]$ is inferred by
          \ruleref{\ottdruleDOneForallLabel}, and by inversion:
          $[[Γ, {pbs}, {pbs'} ⊢ [iPs/pas]iN' ≤ iM']]$ for some $[[Γ, {pbs}, {pbs'} ⊢ iPs]]$.
          Then again by \ruleref{\ottdruleDOneForallLabel} with the same $[[iPs]]$,
          $[[Γ,pbs ⊢ ∀pas.iN' ≤ ∀pbs'.iM']]$, that is $[[Γ,pbs ⊢ iN ≤ iM]]$.
        \item [$\Leftarrow$] let us assume $[[Γ, pbs ⊢ iN ≤ iM]]$, and let us decompose 
          $[[iN]]$ as $[[∀pas.iN']]$ where $[[iN']]$ does not start with $[[∀]]$, 
          and $[[iM]]$ as $[[∀pbs'.iM']]$ where $[[iM']]$ does not start with $[[∀]]$.
          if $[[pas]]$ and $[[pbs']]$ are empty then $[[Γ, pbs ⊢ iN ≤ iM]]$
          is turned into $[[Γ ⊢ iN ≤ ∀pbs.iM]]$ by \ruleref{\ottdruleDOneForallLabel}.
          Otherwise, $[[Γ, pbs ⊢ ∀pas.iN' ≤ ∀pbs'.iM']]$ is inferred by
          \ruleref{\ottdruleDOneForallLabel}, that is $[[Γ, {pbs}, {pbs'} ⊢ [iPs/pas]iN' ≤ iM']]$
          for some $[[Γ, {pbs}, {pbs'} ⊢ iPs]]$.
          Then by \ruleref{\ottdruleDOneForallLabel} again,
          $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs,pbs'.iM']]$, in other words, $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs.∀pbs'.iM']]$, 
          that is $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.

@@

\corollaryRedQuantElim*
\begin{proof}
  \begin{itemize}
    \item [$-_{R}$] Suppose that $[[ {pas} ∩ fv(iM) = ∅]]$ then 
      by \cref{lemma:quant-rule-decomposition},
      $[[Γ ⊢ iN ≤ ∀pas.iM]]$ 
      is equivalent to $[[Γ, pas ⊢ iN ≤ iM]]$,
      By \cref{lemma:wf-ctxt-equiv},
      since $[[{pas} ∩ fv(iN) = ∅]]$ and $[[{pas} ∩ fv(iM) = ∅]]$,
      $[[Γ, pas ⊢ iN ≤ iM]]$ is equivalent to $[[Γ ⊢ iN ≤ iM]]$.

@@

\lemmaVarSubt*
\begin{proof}
  We prove by induction on the tree
  inferring $[[Γ ⊢ iP ≥ ∃nas.α⁺]]$ or $[[Γ ⊢ ∃nas.α⁺ ≥ iP ]]$ or
  or $[[Γ ⊢ iN ≤ ∀pas.α⁻]]$ or $[[Γ ⊢ ∀pas.α⁻ ≤ iN ]]$.

@@

\corollaryVarsNoProperSubtypes*
\begin{proof}
  Notice that $[[Γ ⊢ ∃nbs.α⁺ ≈ α⁺]]$ and $[[∃nbs.α⁺ ≈ α⁺]]$ and apply
  \cref{lemma:var-subt}.
\end{proof}

\lemmaSubtCtxtIrrelevance*
\begin{proof}
  We prove it by induction on 
  the size of $[[Γ1 ⊢ iP ≥ iQ]]$ and mutually,
  the size of $[[Γ1 ⊢ iN ≤ iM]]$.

@@

  Suppose that the judgment is inferred by \ruleref{\ottdruleDOneExistsLabel}.
  Then we are proving that $[[Γ1 ⊢ ∃nas.iP ≥ ∃nbs.iQ]]$ implies 
  $[[Γ2 ⊢ ∃nas.iP ≥ ∃nbs.iQ]]$ (the other implication is proven symmetrically).

@@

  Notice that the premises of the induction hold:
  \begin{enumerate}
    \item $[[Γi, nbs ⊢ [σ0]iP]]$, 
    \item $[[Γi, nbs ⊢ iQ]]$, and
    \item $[[Γ1, nbs ⊢ [σ0]iP ≥ iQ ]]$, 
      notice that the tree inferring this judgment 
      is the same tree inferring $[[Γ1, nbs ⊢ [σ]iP ≥ iQ ]]$ 
      (since $[[ [σ0]iP = [σ]iP]]$), i.e., 
      it is a subtree of $[[Γ1 ⊢ ∃nas.iP ≥ ∃nbs.iQ]]$.
  \end{enumerate}
  This way, by the induction hypothesis,
  $[[Γ2, nbs ⊢ [σ0]iP ≥ iQ ]]$.
  Combining it with $[[ Γ2, nbs ⊢ σ0 :{nas}]]$
  by \ruleref{\ottdruleDOneExistsLabel},
  we obtain $[[Γ2 ⊢ ∃nas.iP ≥ ∃nbs.iQ]]$.

@@

\lemmaSubtWeakening*
\begin{proof}
  By straightforward induction on the subtyping derivation.
  The polymorphic cases follow from \cref{lemma:subst-range-weakening}.
\end{proof}

\lemmaSubtypingReflexivity*
\begin{proof}
  Let us prove it by the size of $[[iN]]$ and mutually, $[[iP]]$.
  \begin{caseof}
    \item $[[iN]] = [[α⁻]]$\\
      Then $[[Γ ⊢ α⁻ ≤ α⁻]]$ is inferred immediately by \ruleref{\ottdruleDOneNVarLabel}.
    \item $[[iN]] = [[∀pas.iN']]$ where $[[pas]]$ is not empty\\
      First, we rename $[[pas]]$ to fresh $[[pbs]]$ in $[[∀pas.iN']]$ to avoid
      name clashes: $[[∀pas.iN']] = [[∀pbs.[pas/pbs]iN']]$.
      Then to infer $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs.[pas/pbs]iN']]$ we can apply 
      \ruleref{\ottdruleDOneForallLabel}, instantiating $[[pas]]$ with $[[pbs]]$:
      \begin{itemize}
        \item $[[fv iN ∩ {pbs} = ∅ ]]$ by choice of $[[pbs]]$,
        \item $[[G, pbs ⊢ pbi]]$,
        \item $[[G, pbs ⊢ [pbs/pas] iN' ≤ [pbs/pas] iN']]$ by the induction hypothesis,
        since the size of $[[ [pbs/pas]iN' ]]$ is equal to the size of $[[iN']]$,
        which is smaller than the size of $[[iN]] = [[∀pas.iN']]$.
      \end{itemize}
    \item $[[iN]] = [[iP → iM]]$\\
      Then $[[Γ ⊢ iP → iM ≤ iP → iM]]$ is inferred by \ruleref{\ottdruleDOneArrowLabel},
      since $[[Γ ⊢ iP ≥ iP]]$ and $[[Γ ⊢ iM ≤ iM]]$ hold the induction hypothesis. 
    \item $[[iN]] = [[↑iP]]$\\
      Then $[[Γ ⊢ ↑iP ≤ ↑iP]]$ is inferred by \ruleref{\ottdruleDOneShiftULabel},
      since $[[Γ ⊢ iP ≥ iP]]$ holds by the induction hypothesis.
    \item The positive cases are symmetric to the negative ones.
  \end{caseof}
\end{proof}

\lemmaSubstPresSubt*
\begin{proof}
  We prove it by induction on the size of the derivation of $[[Γ1 ⊢ iN ≤ iM]]$
  and mutually, $[[Γ1 ⊢ iP ≥ iQ]]$. Let us consider the last rule 
  used in the derivation:
  \begin{caseof}
    \item \ruleref{\ottdruleDOneNVarLabel}. Then by inversion, 
      $[[iN = α⁻]]$ and $[[iM = α⁻]]$. By reflexivity of subtyping
      (\cref{lemma:subtyping-reflexivity}),
      we have $[[Γ2 ⊢ [σ]α⁻ ≤ [σ]α⁻]]$, i.e. $[[Γ2 ⊢ [σ]iN ≤ [σ]iM]]$,
      as required.
    \item  \ruleref{\ottdruleDOneForallLabel}. Then by inversion,
      $[[iN = ∀pas.iN']]$, $[[iM = ∀pbs.iM']]$, where $[[pas]]$ or $[[pbs]]$ is not empty.
      Moreover, $[[Γ1, pbs ⊢ [iPs/pas]iN' ≤ iM']]$ for some $[[Γ1, pbs ⊢ iPs]]$, and 
      $[[fv iN ∩ {pbs} = ∅ ]]$.

@@

\corollarySubstPresEquiv*

\lemmaSubtypingTransitivity*
\begin{proof}
  To prove it, we formulate a stronger property, 
  which will imply the required one, taking $[[σ]] = [[Γ ⊢ id : Γ]]$.

@@

      Notice that the derivations of $[[Γ ⊢ iP ≥ iQ1]]$ and $[[Γ ⊢ iN' ≤ iM1']]$ are proper sub-derivations of
      $[[Γ ⊢ iP → iN' ≤ iQ1 → iM1']]$, and the derivations of $[[Γ ⊢ iQ2 ≥ iR]]$ and $[[Γ ⊢ iM2' ≤ iK']]$ are proper sub-derivations of
      $[[Γ ⊢ iQ2 → iM2' ≤ iR → iK']]$. This way, the induction hypothesis is applicable:
      \begin{itemize}
        \item applying the induction hypothesis to $[[Γ ⊢ iP ≥ iQ1]]$ and $[[Γ ⊢ iQ2 ≥ iR]]$ 
          with $[[Γ' ⊢ σ : Γ]]$ unifying $[[iQ1]]$ and $[[iQ2]]$, we obtain $[[Γ' ⊢ [σ]iP ≥ [σ]iR]]$;
        \item applying the induction hypothesis to $[[Γ ⊢ iN' ≤ iM1']]$ and $[[Γ ⊢ iM2' ≤ iK']]$ 
          with $[[Γ' ⊢ σ : Γ]]$ unifying $[[iM1']]$ and $[[iM2']]$, we obtain $[[Γ' ⊢ [σ]iN' ≤ [σ]iK']]$.
      \end{itemize}
      This way, by \ruleref{\ottdruleDOneArrowLabel}, $[[Γ' ⊢ [σ]iP → [σ]iN' ≤ [σ]iR → [σ]iK']]$,
      that is $[[Γ' ⊢ [σ]iN ≤ [σ]iK]]$, as required.
  \end{caseof}
  After that, we consider all the 
  analogous positive cases and prove them symmetrically.
\end{proof}

\corollaryEquivalenceTransitivity*
